---
title: "Data Analysis - Global Development Index"
output:
  html_document:
    code_folding: "hide"
---

```{r echo=FALSE, message=FALSE, results=FALSE, warning=FALSE, warning=FALSE}
# install packages, if not installed
list.of.packages <- c(
  "reshape2", 
  "plyr", 
  "PerformanceAnalytics",
  "corrplot",
  "Hmisc",
  "hexbin",
  "gapminder",
  "tidyverse",
  "ggplot2",
  "plotly",
  "gganimate"
)


new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

# import packages
lapply(list.of.packages, require, character.only = TRUE)

# read csv
dat.raw <- read.csv('wdi-f.csv')
dat.raw.d <- read.csv('wdi-d.csv')

```

## Introduction
Following report discusses the findings of qunatitative analysis that was conducted on ***World Development Indicators** dataset. In this excercie we tried to study economic, social and enviornmental development of regions and their countires. This document also briefly outlines the walkthrough of process that was followed for this excercise.

The primary aim of this analysis was to study global evolution and to identify factors that contributed in growth or shrikage of countries' perfomance in different sectors, and to summarize findings supported by emperical eveidence.


## Dataset

The dataset used for this analysis was orignally released by Workd Bank as "World Development Indicators". It is a collection of development indicators, compiled from several officially-recognized international sources. The dataset was officialy made available on google big query under public license.

As stated on their official website; 

> ... It (WDI) presents the most current and accurate global development data available, and includes national, regional and global estimates

### Dimensions
Following is the list of dimensions that were obtained as the result of intial pre-processing;

* Indicator
* Region
* Country
* Decade
* Year

### Indicators
Indicators were further narrowed down by grouping them into logical categories. Following are the resulting handpicked indicators;

**a. Social**

 + Pupil-teacher ratio, primary
 + Government expenditure on education, total (% of government expenditure)
 + Access to electricity (% of population)
 + Birth rate, crude (per 1,000 people)
 + Population growth (annual %)
  
**b. Enviornmental**

 + CO2 emissions (metric tons per capita)
 + Total greenhouse gas emissions (kt of CO2 equivalent)
 + Forest area (sq. km)
 + Population density (people per sq. km of land area) 
 + Life expectancy at birth, total (years)

**c. Economic**

 + GDP per capita (current US$)
 + GDP growth (annual %)
 + GDP per capita growth (annual %)
 + Expense (% of GDP)
 + GNI per capita (constant 2010 US$)


## Pre-processing
Final dataset was obtained after running preprocessing internally on R and externally on SQL based platform. Following report only covers pre-processing routine that were covered within R. Queries and code snippets for external pre-processing were hence skipped intentinoally. External pre-prorcessing produced twp CSVs;

* wdi-f.csv
* Group_comparison.csv

Following are code snippets of internal pre-processing;

* Regions were added to all countries

```{r fig.align="center", out.width = '100%'}

#load new data
#dat.new.raw <- read.csv("wdi-e.csv")

# dat.new.raw <- subset(dat.new.raw, select = -c(X, X.2))
# dat.new <- renameIndicators(dat.new.raw)
# dat.newNationalYearlyAvg.w <- groupByCol(dat.new, c('country', 'year', 'indicator'))

# country_year_indicator<-dcast(dat.newNationalYearlyAvg.w, country + year ~ indicator, value.var="value")
# countries <- levels(unique(gapminder$country))
# countries_length <- length(countries)
 
# country_year_indicator$region=""
# for(index in 1:countries_length){
#   country_year_indicator[country_year_indicator$country==countries[index],]$region = as.character(gapminder[gapminder$country==countries[index],]$continent[1][1])
# }
 
# country_year_indicator<-country_year_indicator[country_year_indicator$region!="",]
# write.csv(country_year_indicator, file = "wdi-f.csv")


# label regions; 1=North America, 2=Scandavivan, 3=South Asia, 4=Middle East

labelRegions <- function(dat){
  dat$region[dat$country_code %in% c("USA", "CAN")] <- "North America"
  dat$region[dat$country_code %in% c("DNK", "NOR", "SWE")] <- "Scandinavia"
  dat$region[dat$country_code %in% c("IND", "PAK", "BGD")] <- "South Asia"
  dat$region[dat$country_code %in% c("QAT", "EGY", "ARE")] <- "Middle East"
  
  dat$region <- as.factor(dat$region)
  return(dat)
}

dat.raw.d <- labelRegions(dat.raw.d)

```

* Indicator were reanmed to shorter names for easier referencing

```{r fig.align="center", out.width = '100%'}

renameIndicators <- function(dat){
  
  # old_names
  old_names <- c(
    "GDP per capita (current US$)",
    "GDP growth (annual %)",
    "GDP per capita growth (annual %)",
    "Expense (% of GDP)",
    "Industry (including construction), value added (% of GDP)",
    "Net primary income (BoP, current US$)",
    "GNI per capita (constant 2010 US$)",
    # "GNI per capita (constant LCU)",
    
    "Pupil-teacher ratio, primary",
    "Government expenditure on education, total (% of government expenditure)",
    "Access to electricity (% of population)",
    "Birth rate, crude (per 1,000 people)",
    "Population growth (annual %)",
    "Life expectancy at birth, total (years)",
    
    "CO2 emissions (metric tons per capita)",
    "Total greenhouse gas emissions (kt of CO2 equivalent)",
    "Forest area (sq. km)",
    "Population density (people per sq. km of land area)"
  )
  
  new_names <- c(
    "gdp_capita", # GDP per capita (current US$)
    "gdp_growth", # GDP growth (annual %)
    "gdp_capita_growth", # GDP per capita growth (annual %)
    "expense_gdp", # Expense (% of GDP)
    "industry_gdp", # Industry (including construction), value added (% of GDP)
    "net_primary_income", # "Net primary income (BoP, current US$)"
    "gni_capita", # GNI per capita (constant 2010 US$)
    # "gni_capita_constant", # GNI per capita (constant LCU)
    
    "teacher_per_pupil", # Pupil-teacher ratio, primary
    "education_expense", # Government expenditure on education, total (% of government expenditure)
    "electricity_access", # Access to electricity (% of population)
    "birth_rate", # Birth rate, crude (per 1,000 people)
    "population_growth", # Population growth (annual %)
    "life_expectancy", # Life expectancy at birth, total (years)
    
    "co2_emission", # CO2 emissions (metric tons per capita)
    "greenhouse_emission", # Total greenhouse gas emissions (kt of CO2 equivalent)
    "forest_area", # Forest area (sq. km)
    "population_density" # Population density (people per sq. km of land area)
  )
  
  
  dat$indicator <- mapvalues(dat$indicator, from = old_names, to = new_names)
  
  return(dat)
}

dat.d <- renameIndicators(dat.raw.d)

```

* Dataset was aggregated over different dimensions

```{r fig.align="center", out.width = '100%'}

# group by routine
groupByCol <- function(dat, col, filter = ""){
  groupColumns = col
  dataColumns = c('value')
  
  if(filter != "") {
    
  }
  
  groupedDat <- ddply(
    dat, groupColumns, function(x) colMeans(x[dataColumns])
  )
  return(groupedDat)
}

# add decade column and init gropued by var
dat.d <- mutate(dat.d, decade = year - (year %% 10))


datGlobalYearlyAvg.w <- groupByCol(dat.d, c('year', 'indicator'))
datNationalYearlyAvg.w <- groupByCol(dat.d, c('country', 'year', 'indicator'))
datRegionalYearlyAvg.w <- groupByCol(dat.d, c('region', 'year', 'indicator'))
datNationalDecadeAvg.w <- groupByCol(dat.d, c('country', 'decade', 'indicator'))
datRegionalDecadeAvg.w <- groupByCol(dat.d, c('region', 'decade', 'indicator'))
datRegionalNationalYearlyAvg.w <- groupByCol(dat.d, c('region', 'country', 'year', 'indicator')) 
datRegionalNationalDecadeAvg.w <- groupByCol(dat.d, c('region', 'country', 'decade', 'indicator')) 

datNationalAvg.w <- groupByCol(dat.d, c('country', 'indicator'))
datRegionalAvg.w <- groupByCol(dat.d, c('region', 'indicator'))



```

* Dataset was normalize for some operations using min-max normalization

```{r fig.align="center", out.width = '100%'}

# normalize routine
normalize <- function(col) {
  return((col - min(col))/(diff(range(col))))
}


```

* The orignal structure of dataset was transformed using data transformation techniques to make processing more convenient
* Data was trasnformed from long to wide format

```{r fig.align="center", out.width = '100%'}

# Write CSV in R
#write.csv(dat.all, file = "Final_Data.csv")
change_year_columns <- function (dat){
      colnames(dat)[which(colnames(dat) %in% c("X1960","X1961","X1962","X1963","X1964","X1965","X1966","X1967","X1968","X1969","X1970","X1971","X1972","X1973","X1974","X1975","X1976","X1977","X1978","X1979","X1980","X1981","X1982","X1983","X1984","X1985","X1986","X1987","X1988","X1989","X1990","X1991","X1992","X1993","X1994","X1995","X1996","X1997","X1998","X1999","X2000","X2001","X2002","X2003","X2004","X2005","X2006","X2007","X2008","X2009","X2010","X2011","X2012","X2013","X2014","X2015","X2016","X2017","X2018") )] <- c("1960","1961","1962","1963","1964","1965","1966","1967","1968","1969","1970","1971","1972","1973","1974","1975","1976","1977","1978","1979","1980","1981","1982","1983","1984","1985","1986","1987","1988","1989","1990","1991","1992","1993","1994","1995","1996","1997","1998","1999","2000","2001","2002","2003","2004","2005","2006","2007","2008","2009","2010","2011","2012","2013","2014","2015","2016","2017","2018")
    return (dat)
  }


wide_to_long <- function(data_wide)
{
  data_wide <- change_year_columns(grouped_comparison_wide)

  data_long<-melt(data_wide, id.vars=c("X.1","Country_Name","Country_Code","Indicator_Name","Indicator_Code","X"))
  
  colnames(data_long)[which(colnames(data_long) %in% c("variable") )] <- c("year")
  
  data_long <- data_long[setdiff(colnames(data_long), c('X.1', 'X', "Indicator_Code"))]
  colnames(data_long)[which(colnames(data_long) %in% c("Country_Name","Country_Code","Indicator_Name") )] <- c("country","country_code","indicator")
  return(data_long)
}

datGlobalYearlyAvg <- dcast(datGlobalYearlyAvg.w, year ~ indicator, value.var="value")
datNationalYearlyAvg <- dcast(datNationalYearlyAvg.w, year + country ~ indicator, value.var="value")
datNationalDecadeAvg <- dcast(datNationalDecadeAvg.w, decade + country ~ indicator, value.var="value")
datRegionalYearlyAvg <- dcast(datRegionalYearlyAvg.w, year + region ~ indicator, value.var="value")
datRegionalDecadeAvg <- dcast(datRegionalDecadeAvg.w, decade + region ~ indicator, value.var="value")
datRegionalNationalYearlyAvg <- dcast(datRegionalNationalYearlyAvg.w, year + region + country ~ indicator, value.var="value")
datRegionalNationalDecadeAvg <- dcast(datRegionalNationalDecadeAvg.w, decade + region + country ~ indicator, value.var="value")

datNationalAvg <- dcast(datNationalAvg.w, country ~ indicator, value.var="value")
datRegionalAvg <- dcast(datRegionalAvg.w, region ~ indicator, value.var="value")

```

* Missing value were treated by filling it with mean of respective column

```{r fig.align="center", out.width = '100%'}

fillMissing <- function(df){
  for(i in 1:ncol(df)){
    
    if(is.numeric(df[,i])){
      df[is.na(df[,i]), i] <- mean(df[,i], na.rm = TRUE)
    }
  }
    
  return(df)
}

datGlobalYearlyAvg <- fillMissing(datGlobalYearlyAvg)
datNationalYearlyAvg <- fillMissing(datNationalYearlyAvg)
datNationalDecadeAvg <- fillMissing(datNationalDecadeAvg)
datRegionalYearlyAvg <- fillMissing(datRegionalYearlyAvg)
datRegionalDecadeAvg <- fillMissing(datRegionalDecadeAvg)
datRegionalNationalYearlyAvg <- fillMissing(datRegionalNationalYearlyAvg)
datRegionalNationalDecadeAvg <- fillMissing(datRegionalNationalDecadeAvg)
datNationalAvg <- fillMissing(datNationalAvg)
datRegionalAvg <- fillMissing(datRegionalAvg)


```

* Metric of Social Prosperity and Environment Friendliness was added
* Following were the expressions for above mentioned metrics;

|                **Social Prosperity = log(education_expense*electricity_access*life_expectancy) / log(teacher_per_pupil*birth_rate*population_growth)**

|                **Enviornment Friendliness = log(population_density) / log(co2_emission*greenhouse_emission*forest_area)**


```{r echo=FALSE, message=FALSE, results=FALSE, warning=FALSE, warning=FALSE}

getSocialProsperity <- function(df){
  social_prosperity <- log(df["education_expense"] * df["electricity_access"] * df["life_expectancy"])/ log(df["teacher_per_pupil"] * df["birth_rate"] * df["population_growth"])
  
  return(social_prosperity)
}

getEniviornmenFriendly <- function(df){
  enviornment_friendly <- log(df["population_density"])/ log(df["co2_emission"] * df["greenhouse_emission"]*df["forest_area"])
  
  return(enviornment_friendly)
}


# social prosperity
  datGlobalYearlyAvg['social_prosperity'] <- getSocialProsperity(datGlobalYearlyAvg)
  datRegionalYearlyAvg['social_prosperity'] <- getSocialProsperity(datRegionalYearlyAvg)
  datRegionalDecadeAvg['social_prosperity'] <- getSocialProsperity(datRegionalDecadeAvg)  
  datNationalYearlyAvg['social_prosperity'] <- getSocialProsperity(datNationalYearlyAvg)
  datNationalDecadeAvg['social_prosperity'] <- getSocialProsperity(datNationalYearlyAvg)  
  datNationalAvg['social_prosperity'] <- getSocialProsperity(datNationalAvg)
  datRegionalAvg['social_prosperity'] <- getSocialProsperity(datRegionalAvg)
  datRegionalNationalYearlyAvg['enviornment_friendly'] <- getSocialProsperity(datRegionalNationalYearlyAvg)
  
# enviornemt friendly
  datGlobalYearlyAvg['enviornment_friendly'] <- getEniviornmenFriendly(datGlobalYearlyAvg)  
  datRegionalYearlyAvg['enviornment_friendly'] <- getEniviornmenFriendly(datRegionalYearlyAvg)
  datRegionalDecadeAvg['enviornment_friendly'] <- getEniviornmenFriendly(datRegionalDecadeAvg)  
  datNationalYearlyAvg['enviornment_friendly'] <- getEniviornmenFriendly(datNationalYearlyAvg)
  datNationalDecadeAvg['enviornment_friendly'] <- getEniviornmenFriendly(datNationalDecadeAvg)  
  datNationalAvg['enviornment_friendly'] <- getEniviornmenFriendly(datNationalAvg)
  datRegionalAvg['enviornment_friendly'] <- getEniviornmenFriendly(datRegionalAvg)
  datRegionalNationalYearlyAvg['enviornment_friendly'] <- getEniviornmenFriendly(datRegionalNationalYearlyAvg)
  

```


## Findings

### Economic trends for regions and countries since 1960

```{r fig.align="center", out.width = '100%'}

p <- ggplot(datRegionalYearlyAvg, aes(year, gdp_capita, color = region)) + 
  geom_line() + 
  geom_segment(aes(xstart=1960, xend = 2018, yend = gdp_capita), linetype = 2, colour = 'grey') +
  geom_point(size = 2) + 
  geom_text(aes(x = 2018, label = region), hjust = 0) + 
  transition_reveal(year) + 
  coord_cartesian(clip = 'off') + 
  labs(title = 'National GDP', x='Year', y = 'GDP per Capits') + 
  theme_minimal() + 
  theme(plot.margin = margin(5.5, 40, 5.5, 5.5)) +
  ease_aes('linear', interval = 0.0001)

animate(p, nframes = 300, fps=10)

```

**Regional Economogy**

* Middle east was the leader until 1990
* Scandinavia overtook North America in 1975 and 1995
* South Asia didn’t show any drastic growth
* Disparity started increasing after 1990


```{r fig.align="center", out.width = '100%'}

p <- ggplot(datNationalYearlyAvg, aes(year, gdp_capita, color = country)) + 
  geom_line() + 
  geom_segment(aes(xstart=1960, xend = 2018, yend = gdp_capita), linetype = 2, colour = 'grey') +
  geom_point(size = 2) + 
  geom_text(aes(x = 2018, label = country), hjust = 0) + 
  transition_reveal(year) + 
  coord_cartesian(clip = 'off') + 
  labs(title = 'National GDP', x='Year', y = 'GDP per Capits') + 
  theme_minimal() + 
  theme(plot.margin = margin(5.5, 40, 5.5, 5.5)) +
  ease_aes('linear', interval = 0.0001)

animate(p, nframes = 300, fps=10)

```

**International Economoy**

* UAE and Qatar had highest gdp growths until 1970
* Norway and Denmark showed highest growth in recent decade


### Effect of Social prosperity and enviornmental prosperity over economy, before and after 2000

```{r fig.align="center", out.width = '100%'}

getCorrelation <- function(df){
# slice important cols
  cols <- c(
    "social_prosperity",
    "enviornment_friendly",

    "gdp_growth", # GDP growth (annual %)
    "gni_capita", # GNI per capita (constant 2010 US$)
    # "gni_capita_constant", # GNI per capita (constant LCU)
    
    "education_expense", # Government expenditure on education, total (% of government expenditure)
    "electricity_access", # Access to electricity (% of population)
    "population_growth", # Population growth (annual %)
    "life_expectancy", # Life expectancy at birth, total (years)
    
    "co2_emission", # CO2 emissions (metric tons per capita)
    "forest_area", # Forest area (sq. km)
    "population_density" # Population density (people per sq. km of land area)
  )
  df.slice <- df[,cols]
  
  
  df.slice.corr <- rcorr(as.matrix(df.slice))
  
  return(df.slice.corr)
}

par(mfrow=c(1,2))

# plot 
  # pre-2000
  datGlobalYearlyAvg.slice.corr.1960 <- getCorrelation(datNationalYearlyAvg[datNationalYearlyAvg$year<2000,])
    corrplot(datGlobalYearlyAvg.slice.corr.1960$r, type="upper",
         p.mat = datGlobalYearlyAvg.slice.corr.1960$P, sig.level = 0.01, insig = "blank")
  # 
  # post-2000
  datGlobalYearlyAvg.slice.corr.2018 <- getCorrelation(datNationalYearlyAvg[datNationalYearlyAvg$yea>=2000,])
  corrplot(datGlobalYearlyAvg.slice.corr.2018$r, type="upper",
         p.mat = datGlobalYearlyAvg.slice.corr.2018$P, sig.level = 0.01, insig = "blank")

```

**Before 2000**

* Electricity access decreased, with increase in Population density
* Life Expectancy is heavily dependent on GNI and Social Prosperity

**After 2000**

* All dependencies are still relevant
* Strength of dependencies intensified

### History Over Time: Country's journey to prosperity/failure from 1960 to 2018

+ We were curious to know what is the secret of the successful countries.
+ The report shows the indicators that have the greatest impact on making countries "rich and healthy".
+ We would like to know if we can find the "formula" of the richest and most comfortable countries based on the facts obtained from the indicators.
+ We took y axis for life expectancy and x axis for wealth - Gross National Income Of countries.
+ We wanted to visualise the “movement” of the countries on the graph based on mentioned parameters above. From "poor and sick" up to "rich and healthy".
+ At first, we will show the world 58 years ago in 1960, over the countries of Africa, Asia, Middle East, America and Europe.
+ The size of the point area shows the population density.

```{r fig.align="center", out.width = '100%', message=FALSE, warning=FALSE}

  economic_indicators <- c(
    "gdp_capita", # GDP per capita (current US$)
    "gdp_growth", # GDP growth (annual %)
    "gdp_capita_growth", # GDP per capita growth (annual %)
    "expense_gdp", # Expense (% of GDP)
    "net_income" # GNI per capita (constant LCU)
  )

  social_indicators <- c(
    "teacher_per_pupil", # Pupil-teacher ratio, primary
    "education_expense", # Government expenditure on education, total (% of government expenditure)
    "electricity_access", # Access to electricity (% of population)
    "birth_rate", # Birth rate, crude (per 1,000 people)
    "population_growth" # Population growth (annual %)
  )
  
    environmental_indicators <- c(
    "co2_emission", # CO2 emissions (metric tons per capita)
    "greenhouse_emission", # Total greenhouse gas emissions (kt of CO2 equivalent)
    "forest_area", # Forest area (sq. km)
    "population_density", # Population density (people per sq. km of land area)
    "life_expectancy" #Life expectancy at birth, total (years)
  )
    
read_data <- read.csv('wdi-f.csv')

test <- read_data[FALSE,]

for(index in 1960:2020){
    test <- rbind(test, read_data[read_data$year==index,])
}

gg <- ggplot(test, aes(net_income, life_expectancy, color = region)) +
  geom_point(aes(size = population_density, frame = year, ids = country)) +
  scale_x_log10()+ labs(x = "GNI per capita")

ggplotly(gg)



```


* Colombia / Chile / Costa Rica have a good leap towards higher income
* Benin has a big leap towards life expectancy
* Life expectancy in Bostwana decreased around 1990~2000
* The life expectancy of Ecuador (the poorest country) increases significantly
* Life expectancy in Bangladesh increased to 75 from 45 years over the period (1975-2018)

### Regional population growth trends since 1960

```{r fig.align="center", out.width = '100%'}

# normalize
datRegionalYearlyAvg$population_growth <- normalize(datRegionalYearlyAvg$population_growth)

# extract region
region1 <- datRegionalYearlyAvg[datRegionalYearlyAvg$region=="North America",]
region2 <- datRegionalYearlyAvg[datRegionalYearlyAvg$region=="Scandinavia",]
region3 <- datRegionalYearlyAvg[datRegionalYearlyAvg$region=="South Asia",]
region4 <- datRegionalYearlyAvg[datRegionalYearlyAvg$region=="Middle East",]


# plot
plot(region1$year, region1$population_growth*100, ylim=c(0,100), type = "o", col = 100, pch=19, xlab="Year", ylab = "Growth %", main = "Regional Population Growth")
lines(region2$year, region2$population_growth*100, type = "o", col = 110, pch=20)
lines(region3$year, region3$population_growth*100, type = "o", col = 120, pch=21)
lines(region4$year, region4$population_growth*100, type = "o", col = 130, pch=22)

legend("topright",
  legend = c("North America", "Scandinavia", "South Asia", "Middle East"),
  col = c(100,110, 120,130),
  pch = c(19,20,21,22)
)
```

* Middle east was most populated in history
* North America overtook Scandinavia region in late 2020
* Trends didn’t change throughout history


### Good seprators (indicators) for rich and poor countries

```{r fig.align="center", out.width = '100%', message=FALSE, warning=FALSE}


#Data has been merged before and saved it as a new csv for the sake of simplicity
grouped_comparison_wide <- read.csv('Group_comparison.csv')

grouped_comparison_long <- wide_to_long(grouped_comparison_wide)

grouped_comparison_long <- renameIndicators(grouped_comparison_long)

grouped_comparison_long<-grouped_comparison_long[!is.na(grouped_comparison_long$value),]

grouped_comparison_indicators_wide <- dcast(grouped_comparison_long, year+country ~ indicator, value.var="value")
#grouped_comparison_indicators_wide[, -c(1,2)] <- scale(grouped_comparison_indicators_wide[, -c(1,2)])


long<-melt(grouped_comparison_indicators_wide, id.vars=c("year","country"))

llong<-within(long, {
    group = ifelse(variable %in% social_indicators, "Social", "Environmental")
    group = ifelse(variable %in% environmental_indicators, "Environmental", "Economical")
})

check_val <- function(x) {
  #print(x)
  if(x%in% social_indicators) {
    "Social"
    #print("Social")
  }
  else if(x%in% environmental_indicators) {
    "Environmental"
    #print("Environmental")
  }
  else if(x%in% economic_indicators) {
    "Economical"
    #print("Economical")
  }
}

llong$group <- apply(array(long[["variable"]]),MARGIN=1, FUN=check_val)
llong <- llong[!is.na(llong$value),]
#llong<-llong[llong$year<-2008,]
llong$year <-as.numeric(as.character(llong[,1])) 


```


```{r fig.align="center", out.width = '100%', message=FALSE, warning=FALSE}

plot_circular_bar<-function(llong,var1,var2){
  
  # library
  library(tidyverse)
  llong<-llong[llong$group==var1,]
  llong<-llong[llong$country==var2,]
  #view(llong)
  llong$variable<-factor(llong$variable)
  #print( llong$variable) 
  # Create dataset
  if (var1=="Environmental")
  {
    for (i in 1:length (environmental_indicators)){
      #print(environmental_indicators[i])
      llong[llong$variable==environmental_indicators[i],]$value = normalize(llong[llong$variable==environmental_indicators[i],]$value)
      
    }
      data=data.frame(
        individual=llong$year,
        group=llong$variable ,
        value=llong$value*50)
  }
  else if (var1=="Economical"){
    
        for (i in 1:length (economic_indicators)){
      #print(environmental_indicators[i])
      llong[llong$variable==economic_indicators[i],]$value = normalize(llong[llong$variable==economic_indicators[i],]$value)
      
    }
    
      data=data.frame(
        individual=llong$year,
        group=llong$variable ,
        value=llong$value*50)
  }
  else{
            for (i in 1:length (social_indicators)){
      #print(environmental_indicators[i])
      llong[llong$variable==social_indicators[i],]$value = normalize(llong[llong$variable==social_indicators[i],]$value)
      
    }
    
      data=data.frame(
        individual=llong$year,
        group=llong$variable ,
        value=llong$value*50)
  }

  #data = data %>% arrange(group, value)    
  
  # Set a number of 'empty bar' to add at the end of each group
  empty_bar=2
  to_add = data.frame( matrix(NA, empty_bar*nlevels(data$group), ncol(data)) )
  colnames(to_add) = colnames(data)
  to_add$group=rep(levels(data$group), each=empty_bar)
  data=rbind(data, to_add)
  data=data %>% arrange(group)
  data$id=seq(1, nrow(data))
  
  # Get the name and the y position of each label
  label_data=data
  number_of_bar=nrow(label_data)
  angle= 90 - 360 * (label_data$id-0.5) /number_of_bar     # I substract 0.5 because the letter must have the angle of the center of the bars. Not extreme right(1) or extreme left (0)
  label_data$hjust<-ifelse( angle < -90, 1, 0)
  label_data$angle<-ifelse(angle < -90, angle+180, angle)
  
  # prepare a data frame for base lines
  base_data=data %>% 
    group_by(group) %>% 
    summarize(start=min(id), end=max(id) - empty_bar) %>% 
    rowwise() %>% 
    mutate(title=mean(c(start, end)))
  
  # prepare a data frame for grid (scales)
  grid_data = base_data
  grid_data$end = grid_data$end[ c( nrow(grid_data), 1:nrow(grid_data)-1)] + 1
  grid_data$start = grid_data$start - 1
  grid_data=grid_data[-1,]
  
  
  # Make the plot
  p = ggplot(data, aes(x=as.factor(id), y=value, fill=group)) +       # Note that id is a factor. If x is numeric, there is some space between the first bar
    
    geom_bar(aes(x=as.factor(id), y=value, fill=group), stat="identity", alpha=0.5) +
    
    # Add a val=100/75/50/25 lines. I do it at the beginning to make sur barplots are OVER it.
    geom_segment(data=grid_data, aes(x = end, y = 80, xend = start, yend = 80), colour = "grey", alpha=1, size=0.3 , inherit.aes = FALSE ) +
    geom_segment(data=grid_data, aes(x = end, y = 60, xend = start, yend = 60), colour = "grey", alpha=1, size=0.3 , inherit.aes = FALSE ) +
    geom_segment(data=grid_data, aes(x = end, y = 40, xend = start, yend = 40), colour = "grey", alpha=1, size=0.3 , inherit.aes = FALSE ) +
    geom_segment(data=grid_data, aes(x = end, y = 20, xend = start, yend = 20), colour = "grey", alpha=1, size=0.3 , inherit.aes = FALSE ) +
    
    # Add text showing the value of each 100/75/50/25 lines
    annotate("text", x = rep(max(data$id),4), y = c(20, 40, 60, 80), label = c("20", "40", "60", "80") , color="grey", size=1 , angle=0, fontface="bold", hjust=1) +
    
    geom_bar(aes(x=as.factor(id), y=value, fill=group), stat="identity", alpha=0.5) +
    ylim(-100,120) +
    theme_minimal() +
    theme(
      legend.position = "right",
      axis.title = element_blank(),
      axis.text = element_blank(),
      panel.grid = element_blank(),
      plot.margin = unit(rep(-1,4), "cm") 
    ) +
    coord_polar() + 
    geom_text(data=label_data, aes(x=id, y=value+60, label=individual, hjust=hjust), color="black",alpha=0.6, size=2.5, angle= label_data$angle, inherit.aes = FALSE ) +
    
    # Add base line information
    geom_segment(data=base_data, aes(x = start, y = -5, xend = end, yend = -5), colour = "black", alpha=0.8, size=0.6 , inherit.aes = FALSE )
  
  #   if (var1=="Economical")
  #   {
  #       #print(var1)
  #       #print(var2)
  #       if(var2=="Low income"){
  #         geom_text(data=base_data, aes(x = title, y = -18, label=group), hjust=c(1,1,0), colour = "black", alpha=0.8, size=2, fontface="bold", inherit.aes = FALSE)
  #       }
  #         else
  #       {
  #                   geom_text(data=base_data, aes(x = title, y = -18, label=group), hjust=c(1,1,1,0), colour = "black", alpha=0.8, size=2, fontface="bold", inherit.aes = FALSE)
  #       }
  #   }
  # else
  # {
  #         #print(base_data)
  #         geom_text(data=base_data[1:5,], aes(x = title, y = -18, label=group), hjust=c(1,1,1,0,1), colour = "black", alpha=0.8, size=3, fontface="bold", inherit.aes = FALSE)
  # }
 if(var1=="Economical" && var2=="Low income")
 {
   #print(paste(var2,":",var1))
   #"#9ACD32", "#00CED1", "#BA55D3"
   p + scale_fill_manual(paste(var2,":",var1), values = c("gdp_growth" = "#9ACD32", "gdp_capita" = "#00CED1", "gdp_capita_growth" = "#BA55D3"))
 }
  else{
     p + scale_fill_discrete(name = paste(var2,":",var1))
  }
  
  
  #print(data$group)
}


```

#### Social indicators and High/Low/Mid income countries

**a. High Income countries**

```{r fig.align="center", out.width = '100%', message=FALSE, warning=FALSE}

plot_circular_bar(llong,"Social","High income")

```

+ Pupil-teacher ratio, primary: Decreases with time. Lower after 2010.
+ Government expenditure on education, total (% of government expenditure): Not consistent, but maintains certain threshold.
+ Access to electricity (% of population): Increases with time.
+ Birth rate, crude (per 1,000 people): Decreases with time.
+ Population growth (annual %): Decreases with time. After 2010 almost constant.

**b. Middle Income countries**

```{r fig.align="center", out.width = '100%', message=FALSE, warning=FALSE}

plot_circular_bar(llong,"Social","Middle income")

```

+ Pupil-teacher ratio, primary: Higher compared to rich countries and increases more.
+ Government expenditure on education, total (% of government expenditure): Increases but we have less data here.
+ Access to electricity (% of population): Almost equal and increases over time.
+ Birth rate, crude (per 1,000 people): A bit Higher but also decreases more.
+ Population growth (annual %): Decreases more compared to High income countries.

Pupil-teacher ratio, Birthrate are the indicators which are separating Middle from High income countries.


**c. Low Income countries:**

```{r fig.align="center", out.width = '100%', message=FALSE, warning=FALSE}

plot_circular_bar(llong,"Social","Low income")

```

+ Pupil-teacher ratio, primary: Trend is totally opposite compared to above groups(inconsistent).
+ Government expenditure on education, total (% of government expenditure): Decreases with time which is not ideal compared to other groups.
+ Access to electricity (% of population): Higher which is similar to high and middle income countries.
+ Birth rate, crude (per 1,000 people): Decreases which is also a good indicator.
+ Population growth (annual %): Not consistent and also higher.

Pupil-teacher ratio, primary, Government expenditure on education and population growth are the clear distinction between High and Middle income countries.


#### Environmental indicators and High/low/mid income countries

**a. High Income countries**

```{r fig.align="center", out.width = '100%', message=FALSE, warning=FALSE}

plot_circular_bar(llong,"Environmental","High income")

```

+ CO2 emissions (metric tons per capita): Maintains threshold value after 1967.
+ Total greenhouse gas emissions (kt of CO2 equivalent): Increases with time
+ Forest area (sq. km): Increases with time.
+ Population density (people per sq. km of land area): Increase with time.
+ Life expectancy at birth, total (years): Increases with time
 

**b. Middle Income countries**

```{r fig.align="center", out.width = '100%', message=FALSE, warning=FALSE}

plot_circular_bar(llong,"Environmental","Middle income")

```

+ CO2 emissions (metric tons per capita): Lower compared to rich countries and increases.
+ Total greenhouse gas emissions (kt of CO2 equivalent): Lower but increasing.
+ Forest area (sq. km): Decreasing over time (Opposite trend spotted).
+ Population density (people per sq. km of land area): Increases almost the same as high income.
+ Life expectancy at birth, total (years): Increases almost the same as high income.

CO2 emissions, Forest area and Greenhouse emission  are in some sense good distinguishing factors.


**c. Low Income countries**

```{r fig.align="center", out.width = '100%', message=FALSE, warning=FALSE}

plot_circular_bar(llong,"Environmental","Low income")


```

+ CO2 emissions (metric tons per capita): Much higher, almost the same as in High income countries.
+ Total greenhouse gas emissions (kt of CO2 equivalent): Increased after 2000 and almost did not change.
+ Forest area (sq. km): Higher compared with the Middle income, but a similar trend.
+ Population density (people per sq. km of land area): Lower than in Middle income countries, but also increases with time.
+ Life expectancy at birth, total (years): Lower than Middle income countries and increases over time.

High and Low income countries make a larger contribution to environmental pollution.


#### Economical indicators and High/low/mid income countries

**a. High Income countries**

```{r fig.align="center", out.width = '100%', message=FALSE, warning=FALSE}

plot_circular_bar(llong,"Economical","High income")


```

+ GDP per capita (current US$): Increases with time.
+ GDP growth (annual %): Increases over time.
+ GDP per capita growth (annual %): No obvious pattern but above a certain threshold.
+ Expense (% of GDP): Increases with time.


**b. Mid Income countries**

```{r fig.align="center", out.width = '100%', message=FALSE, warning=FALSE}

plot_circular_bar(llong,"Economical","Middle income")


```

+ GDP per capita (current US$): Probably lower and also increases with time
+ GDP growth (annual %): Lower and no obvious pattern over time
+ GDP per capita growth (annual %): Lower and also no obvious pattern
+ Expense (% of GDP): Low data

**c. Low Income countries**

```{r fig.align="center", out.width = '100%', message=FALSE, warning=FALSE}

plot_circular_bar(llong,"Economical","Low income")

```

+ GDP per capita (current US$): Lower and increases with time
+ GDP growth (annual %): Probably lower and no clear pattern but maintains threshold.
+ GDP per capita growth (annual %): Probably lower. No clear pattern but maintains a threshold.
+ Expense (% of GDP): No data.
+ GNI per capita (constant LCU): No data.

All these groups have a similar tendency of growth of GDP per capita, but clear distinction could be spotted on GDP and GDP per capita growth.
High and Middle income countries has higher growth.

#### The challenges we faced:
The “Circular barplot with groups” were used to visualise the difference between Hight, Middle and Low income countries by 3 indicators. Each bar represents a value. The problem we faced was that some indicators have missing values (NA) in particular years. And if we plotted the data without removing the missing values we get so much “empty” bars (with no value meaning) that it impossible to interpret the data. That is why it was necessary to take out the missing values, in order to have visible and understandable representation.


### Effect of social prosperity on economic prosperity

```{r}
ggplot(datRegionalDecadeAvg, aes(decade, social_prosperity, fill = social_prosperity)) +
  geom_col() +
  scale_fill_distiller(palette = "Reds", direction = 1) +
  facet_wrap(~region) +
  labs(x = 'Decade', y = 'Social Prosperity Score') +
  transition_states(decade) +
  shadow_mark()
```

* Economic success resulted social prosperity for region North America and Scandinavia
* Social prosperity of region South Asia and Middle East has not witness any drastic change

## References
https://datacatalog.worldbank.org/dataset/world-development-indicators
http://www.humantruth.info/regional_comparisons.html
https://www.sciencedirect.com/topics/social-sciences/socio-economic-indicators http://www.fao.org/3/y5228e/y5228e03.htm
https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf